### 2011-07-26 Introduction.tex
Filling in the images

\section {Introduction}

%The chemistry of our atmosphere is vastly complex and poorly understood. Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations. How do atmospheric conditions affect the chemistry of environmental constituents?