### 2011-08-31 Article.tex
old?

%\documentclass{article}
\documentclass{scrartcl}

\usepackage {amsmath}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
\usepackage{lineno}

% For typesetting roman numerals
\makeatletter
\newcommand{\rmnum}[1]{\romannumeral #1}
\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

% some formatting tags
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Further Explorations of \suldiox~Aqueous Adsorption:}
\subtitle{Coordinated Bonding and Cyclic Hydrates}
%\author{Eric S. Shamay \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

%\maketitle

%\onehalfspacing
\linenumbers 
\doublespacing

%\begin{abstract}
	%...Abstract goes here...
%\end{abstract}

\include{Introduction}
\include{Background}
\include{ComputationalMethods}
\include{Coordination-Results}
\include{Cycle-Results}
%\include{orientation}
%\include{sfg}
%\include{Conclusions}

\bibliography{bib/so2temperature}

\end{document}