### 2011-12-05 Article.tex
Draft 3 sent to Geri, along with new abstract and title.

\documentclass{article}
%\documentclass{scrartcl}

\usepackage {amsmath}
\usepackage {textcomp}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
%\usepackage{lineno}

% For typesetting roman numerals
\makeatletter
\newcommand{\rmnum}[1]{\romannumeral #1}
\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

% some formatting tags
\usepackage[
top    = 0.5in,
bottom = 1.5in,
left   = 2.50cm,
right  = 2.50cm]{geometry}
%\oddsidemargin  0.0in
%\evensidemargin 0.0in
%\topmargin -0.5in
%\bottommargin -0.5in
%\textwidth      6.5in

\title{Bonding, Intimacy, and Circles of Friends: Further Computational Studies of Sulfur Dioxide Adsorption to Aqueous Surfaces}
\author{Eric S. Shamay \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

\maketitle

%\onehalfspacing
%\linenumbers 
\doublespacing

\begin{abstract}
A water surface is a dynamic and constantly evolving terrain producing a vast array of unique molecular properties and interactions with environmental chemical species. The dynamics of water surfaces permit life on earth to continue, but also keeps us from developing a complete microscopic picture of the specific behaviors that take place within interfacial aqueous environments. This computational study examines a piece of the water puzzle by elucidating the bonding, dynamic interactions, and hydrate structures of sulfur dioxide gas adsorbing to a water surface. Results described herein address the specific ways in which sulfur dioxide gas molecules bind to a water surface, and paint a more complete picture of the adsorption pathway than was previously developed from experimental and computational studies. Ab initio molecular dynamics have been employed to study sulfur dioxide and water interactions at two environmentally relevant temperatures on a water surface. The results of this study on a common environmental and industrially important gas provide molecular insight to aid our understanding of interactions on aqueous surfaces, and gaseous adsorption processes.
\end{abstract}

\include{Introduction}
\include{Background}
\include{ComputationalMethods}
\include{Coordination-Results}
\include{Cycle-Results}
\include{Conclusions}

\bibliography{bib/so2temperature}

\end{document}